Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-13 chapters/LBM_technique.tex
Changes 1. Added paper attribution as footnotes 2. Fixed some language in lit review

>>>>

\section{Simulation setup}
\label{section:sim_setup}

Critical parameters in this system include fluid density $\rho_f$, dynamic viscosity $\eta$, surface tension $\sigma$, 
length scale $L$, and applied magnetic field strength $B$. The dynamic viscosity and surface tension is determined 
by the relaxation time using the BGK collision operator, $\eta = \rho_f c_s^2(\tau - \frac{1}{2})$ and $g_{kk'}$ 
respectively. Dimensionless numbers derived from the Buckingham Pi theorem are the Reynolds number
($Re = \frac{\rho u L}{\eta}$), Weber number ($We = \frac{\rho u^2 L}{\sigma}$), and magnetic bond number 
($\Bar{B} = \frac{Bd}{\sigma A}$). Here, $u$ is velocity, $d$ is the particle's dipole moment, and $A$ is the 
particle's cross-sectional area. The system's velocity and length scale are set as the domain size and coarsening rate. 
Parameters for the fluid, particle, and non-ideal mixing model used in the simulations are summarized in Table 
\ref{table:model_params}.

\begin{table}[h!]
\centering
\begin{tabular}{||c c c c c c c c c c c c c||} 
 \hline
 $t_{sim}$ & $L_x$ & $\rho_f$ & $\tau$ & $\eta_f$ & $g_{cc'}$ & $\sigma$ & $\phi_f$ & $\rho_p$ & $\phi_p$ & $V_p$ & $d_c$ & $K_H$ \\ [0.5ex] 
 \hline\hline
 $10^5$ & 256 & 0.7 & 1 & 0.117 & 0.08 & 0.0267 & 0.5 & 1 & 0.15 & 2000$\pi$ & 2/3 & 100\\ [1ex] 
 \hline
\end{tabular}
\caption{Summary of the fluid, particle and model parameters used in the proposed simulations}
\label{table:model_params}
\end{table}

In these simulations, the aspect ratio is defined as $\alpha = \frac{R_{\parallel}}{R_{\perp}}$ with $\alpha = 0.5, 1, 2$ 
used in these simulations. All particle geometries are volume matched, based on the spherical particle volume. For the 
spherical particle, the radius was set at $R_{\parallel} = R_{\perp} = 7.9$ which sets the $\alpha = 0.5$ particles to 
$R_{\parallel} = 5, R_{\perp} = 10$ and the $\alpha = 2$ particles to $R_{\parallel} = 12.6, R_{\perp} = 6.3$. 
$R_{\parallel}$ and $R_{\perp}$ is the radius of the particle parallel and perpendicular to the symmetry axis of 
the particle respectively. These values are identical to what was used in Gunther et al, allowing comparisons to the 
trends they observed. \cite{gunther_timescales_2014} The particles were randomly placed in the system and an equilibration
step was performed to push the particles further away from each after placement. Once equilibration is performed, a 
constant magnetic field is switched on with dimensionless strengths defined using $\bar{B}$.

The initial condition of the system when simulating bijels sets the density of each grid to $0.5\rho_f + \varepsilon$ of 
each fluid where $\varepsilon = 0.0001N(0,1)$. The model parameters trigger phase separation with a sharp interface at 
the start of the simulation. Periodic boundary conditions are applied on all sides unless the system is undergoing shear. 
The shear gradient is applied as a velocity in the z direction across the x axis and implemented as a Lees-Edwards 
(LE) boundary condition in the yz planes. \cite{wagner_leesedwards_2002, lorenz_lees-edwards_2009, yang_capillary_2022} 
These boundaries move with velocities $u_{LE}$ and $-u_{LE}$ to conserve momentum, with a shear rate 
$\dot{\gamma} = \frac{2 u_{LE}}{L_x}$. LE boundaries are preferred over walls to avoid affecting viscosity measurements. 
\cite{wagner_leesedwards_2002, lorenz_lees-edwards_2009, yang_capillary_2022}

Bijel templates have been synthesized using water/2-6-lutidine for further applications. \cite{lee_making_2013} 
While intrinsically polymerizable bijels have been prepared using oligomeric polymers, STRIPS and VIPS currently 
utilize small molecule liquids in their bijel casting mixtures. Therefore, an experimentally viable system using 
water/2-6-lutidine will be discussed in the context of the work presented here. The critical temperature (UCST) 
for this system is $T_c = 34 ^{\circ}C$, with many systems heated to $T = 40^{\circ}C$ to form the bijel. At this 
temperature, the fluid properties are $\sigma = 0.22 mN/m$ and $\eta = 2.38 mPas$. \cite{grattoni_lower_1993} 
Fei et al. developed a method to synthesize polystyrene microparticles ($(R_p = 2 \mu m)$) with tunable geometries 
by mechanically stretching or compressing them, then coating them with nickel to create magnetic microparticles with 
a dipole moment of approximately$\approx 3 \cdot 10^{-14} Am^2$. \cite{fei_active_2017, fei_magneto-capillary_2020} 
Comparing the physical properties to simulation parameters, the simulations correspond to a length scale of
$\Delta x = 252 \cdot 10^{-9} \text{m}$, timescale of $\Delta t = 624 \cdot 10^{-9} \text{s}$, 
mass scale of $\Delta m = 2.24 \cdot 10^{-15} \text{kg}$ and charge scale of $\Delta q = 2.948 \cdot 10^{-7} \text{C}$.

\section{Model results}
\label{section:model_results}

\subsection{Surface tension}
\label{section:model_surface_tension}

In experiments, Wilhelmy and ring tensionometry are used to measure the surface tension from the force needed 
to break through the interface. Pendant drop tensionometry is also used by measuring the pressure required to 
maintain a droplet of analyte at a various sizes and utilizing the Young-Laplace equation to calculate the 
surface tension of the system. \cite{sun_assembly_2013, berry_measurement_2015}
Upon extrusion of a droplet from a reservoir, the size of a droplet is determined by the balance between 
cohesive forces between bubble molecules and geometrical forces that attempt to reduce the curvature of a 
system by increasing its size. For spherical droplets, the equation simplifies to 
$\Delta P = \frac{2 \sigma}{R_d}$ where $\Delta P$ is the difference in scalar pressure 
between the center of the droplet and the ambient pressure, $\sigma$ is the surface tension 
of the interface and $R_d$ is the radius of the droplet.

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.5]{figures/model_validation/surface_tension.png}
    \caption{Plot of the surface tension of the multicomponent model in this work calculated using the 
    Young-Laplace equation.}
    \label{fig:young_laplace_valid}
\end{figure}

Using the fluid and non-ideal mixing parameters defined in Table \ref{table:model_params}, a single water in oil 
droplet, was initialized in the center of the system and run till the droplet radii reached steady state. 
This process was repeated for various droplet diameters, specified between $70$ and $85 \%$ of the system size at $5\%$ 
increments. The system had a box length of $L = 64$ and was run for $50000$ timesteps with periodic boundary conditions 
on all sides. Data was dumped every 10000 timesteps with only the last timestep used to calculate the values shown. Due 
to the diffuse interface present in this method, the density profiles were verified to be at steady state at the end of 
the simulation before further analysis was performed. \cite{frijters_effects_2012} Past work also showed that the 
pressures remained constant after 5 lattice units away from the interface. \cite{frijters_effects_2012}

The pressure difference was calculated using the difference in scalar pressure between the center of the droplet and 
one of the corners of the box. Two techniques exist to calculate the radius of the droplet. The first directly measures 
the distance between opposite points of the droplet as measured from the density profile at opposite points of the droplet, 
and the second is based upon the total mass and density of the fluid making up the droplet. The second method offers greater 
accuracy by avoiding discretization errors that may exist in the first method, and also allows for the correction of 
particle volume at the interface if the system contained them. First, a local effective mass density 
$\rho^b_{eff} = \rho^{b}_{d} - \rho^{b}_{m}$ is found by calculating the difference in density between the blue 
fluid in the droplet and in the matrix. Next, the droplet mass is calculated as $M_d = \sum_{\mathbf{x}}{\rho_{eff}^{b}}$. 
The radius of the droplet can then be calculated from the mass and density of the droplets as $R_d = \sqrt[3]{\frac{3}{4\pi} 
\frac{M_d}{\rho^b_d - \rho^b_m}}$. To linearize the plot, the $\Delta P$ was plotted against $\frac{2}{R_d}$ with 
$\sigma$ being the slope of the fit.

Figure \ref{fig:young_laplace_valid} shows the results of the fit conducted of the Young-Laplace equation demonstrating 
that under these fluid and mixing parameters, the surface tension is $\sigma = 0.0267$. The fit was performed using 
\texttt{scipy.optimize.curve\_fit} with the fit function defined as $y = mx$.

### 2025-04-13 chapters/lit_review_new.tex
Changes 1. Added paper attribution as footnotes 2. Fixed some language in lit review

At the binodal, the system is still metastable, meaning that unless perturbations are introduced to form nuclei, the mixture remains homogeneous \cite{thanh_mechanisms_2014}.
However, a deeper quench into the two-phase region leads to the spinodal line, defined by the stability criterion \(\frac{d^2G}{dx^2} = 0\) \cite{cahn_free_1958}. 
Within this region, the system is unstable to even infinitesimal composition fluctuations, and phase 
separation proceeds spontaneously without the need for nucleation. This process, known as spinodal decomposition, is characterized by the continuous and simultaneous growth of interconnected 
domains of both phases \cite{cahn_free_1958}. Unlike nucleation and growth, which relies on localized events, spinodal decomposition 
results in a bicontinuous structure driven by the system's attempt to minimize 
free energy through composition modulation \cite{cahn_free_1958}. The dynamics of spinodal decomposition were characterized by Cahn and Hilliard \cite{cahn_spinodal_1961}. 
We first begin by defining the free energy of a regular solution given by,

\begin{equation}
    G_m (c) = \chi (1 - c)c + N_a k_B T \left[(1 - c)\ln(1 - c) + c \ln(c)\right]
\end{equation}

This bulk free energy characterizes the thermodynamic driving forces for phase separation. Alternative forms of the free energy can be defined based on the double well potential or
Flory-Huggins theory based on the desired objectives of the model. In real systems, the interface between two partially miscible fluids is diffuse, meaning that the composition 
transitions smoothly across a finite interface width. To account for this, a gradient energy penalty is added to the free energy to penalize sharp spatial variations in 
composition, yielding the free energy functional of the system. The total free energy functional is given by,

@@

where $\kappa$ is the gradient energy coefficient controlling the interface width. Larger \(\kappa\) values correspond to wider interfaces.
The chemical potential is then defined as the variational derivative of the free energy functional,

>>>>

At the binodal, the system is still metastable, meaning that unless perturbations are introduced to form nuclei, the mixture remains homogeneous \cite{thanh_mechanisms_2014}.
However, a deeper quench into the two-phase region leads to the spinodal line, defined by the stability criterion \(\frac{d^2G}{dx^2} = 0\) \cite{cahn_free_1958}. 
Within this region, the system is unstable to even infinitesimal composition fluctuations, and phase 
separation proceeds spontaneously without the need for nucleation. This process, known as spinodal decomposition, is characterized by the continuous and simultaneous growth of interconnected 
domains of both phases \cite{cahn_free_1958}. Unlike nucleation and growth, which relies on localized events, spinodal decomposition 
results in a bicontinuous structure driven by the system's attempt to minimize 
free energy through composition modulation \cite{cahn_free_1958}. The dynamics of spinodal decomposition were characterized by Cahn and Hilliard \cite{cahn_spinodal_1961}. 
We first begin by defining the free energy of a regular binary solution given by,

\begin{equation}
    \Delta G_m (c) = \chi (1 - c)c + N_a k_B T \left[(1 - c)\ln(1 - c) + c \ln(c)\right]
\end{equation}

When $\chi = 0$ the free energy of an ideal solution is recovered. When $\chi > 0$ repulsive interactions between the components 
causes the system to phase separate. Alternative forms of the free energy can be defined based on the double well potential or
Flory-Huggins theory based on the desired objectives of the model. In real systems, the interface between two partially miscible fluids is 
diffuse, meaning that the composition transitions smoothly across a finite interface width. To account for this, a gradient energy penalty is 
added to the free energy to penalize sharp spatial variations in composition, yielding the free energy functional of the system. 
The total free energy functional is given by,

@@

where $\kappa$ is the gradient energy coefficient controlling the interface width. Larger \(\kappa\) values correspond to wider interfaces.
$f(c)$ is obtained from $\Delta G_m (c)$.
The chemical potential is then defined as the variational derivative of the free energy functional,

### 2025-04-13 chapters/paper1_5.tex
Changes 1. Added paper attribution as footnotes 2. Fixed some language in lit review

\section{Introduction}

This section is based on work submited to the American Institude of Physics, Physics of Fluids journal.
It is reproduced from submission number \#POF25-AR-DSFD2024-03180, with the permission of AIP Publishing.

@@

\section{Results}

@@

\section{Conclusions\label{conclusions}}

>>>>

\section[Introduction]{Introduction\protect\footnote{This section is based on work submited to the American Institude of Physics, Physics of Fluids journal.
                        It is reproduced from submission number \#POF25-AR-DSFD2024-03180, with the permission of AIP Publishing.}}

@@

\section{Results}
% \section[Results]{Results\protect\footnote{This section is based on work submited to the American Institude of Physics, Physics of Fluids journal.
%                     It is reproduced from submission number \#POF25-AR-DSFD2024-03180, with the permission of AIP Publishing.}}

@@

\section{Conclusion}
% \section[Conclusion]{Conclusions\protect\footnote{This section is based on work submited to the American Institude of Physics, Physics of Fluids journal.
% It is reproduced from submission number \#POF25-AR-DSFD2024-03180, with the permission of AIP Publishing.}}

### 2025-04-13 chapters/paper1_new.tex
adjusted footnote
Changes 1. Updating some refereces of analysis to link previous sections which do the same analysis to prevent repetition 2. Updating paper 2 to remove junk results
Changes 1. Added paper attribution as footnotes 2. Fixed some language in lit review

\section{Introduction}

This section is based on work published in the Royal Society of Chemistry's Soft Matter journal. \cite{karthikeyan_formation_2024} 
It is reproduced from Ref \cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.

@@

\section{Results}

@@

    In LB simulations of spinodal decomposition
    \cite{kendon_3d_1999,kendon_inertial_2001}, a characteristic length
    scale can be obtained from the structure factor
    %
    \begin{equation}
    S(\vec{k},t) = \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) .
    \end{equation}
    %
    where \(\tilde{\phi}(\vec{k},t)\) is the Fourier
    transform of the order parameter fluctuations
    \(\phi(\vec{x},t)-\left\langle\phi\right\rangle\). A common measure of
    domain size uses the spherically averaged structure factor
    %
    \begin{equation}
    S(k,t) = \frac{1}{n_k} \sum_{k-\frac{\Delta}{2}<k<k+\frac{\Delta}{2}} \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) ,
    \end{equation}
    %
    where \(k=|\vec{k}|\) denotes the modulus of \(\vec{k}\)
    and \(n_k\) is the number of lattice sites in a shell of radius \(k\)
    and thickness \(\Delta=2\pi/L_V\). The average domain size can then be
    defined in terms of the \(n\)-th moment of the structure factor
    \cite{laradji_molecular_1996}
    %
    \begin{equation}
        L_n(t) = 2\pi \left( \frac{\sum_k S(k,t)}{\sum_k k^n S(k,t)} \right)^{\frac{1}{n}}
    \end{equation}

@@

    %
    \begin{equation}
    L_\beta(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}k_\beta^2 S(\vec{k},t)}} .
    \end{equation}
    %
    This allows us to compute a domain size parallel
    (\(L_\parallel=L_z\)) and perpendicular (\(L_\perp=(L_x+L_y)/2\)) to the
    direction of \(\vec{B}\) separately. The average domain size \(L_d(t)\)
    is computed as \(L_d(t)=\sum_\beta L_\beta(t) / 3\).
    %
    \begin{equation}
    L_d(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}\sum_\beta k_\beta^2 S(\vec{k},t)}} .
    \end{equation}
    %

@@

\section{Conclusions}

>>>>

% \section{Introduction}

% This section is based on work published in the Royal Society of Chemistry's Soft Matter journal.
% It is reproduced from Ref \cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.

\section[Introduction]{Introduction\protect\footnote{This section is based on work published in the Royal Society of Chemistry's Soft Matter journal. It is reproduced from Ref.~\cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.}}

@@

% \section[Results]{Results\protect\footnote{This section is based on work published in the Royal Society of Chemistry's Soft Matter journal. It is reproduced from Ref.~\cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.}}
\section{Results}

@@

    In LB simulations of spinodal decomposition
    \cite{kendon_3d_1999,kendon_inertial_2001}, a characteristic length
    scale can be obtained from the structure factor
    %
    \begin{equation}
    S(\vec{k},t) = \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) .
    \end{equation}
    %
    where \(\tilde{\phi}(\vec{k},t)\) is the Fourier
    transform of the order parameter fluctuations
    \(\phi(\vec{x},t)-\left\langle\phi\right\rangle\). A common measure of
    domain size uses the spherically averaged structure factor
    %
    \begin{equation}
    S(k,t) = \frac{1}{n_k} \sum_{k-\frac{\Delta}{2}<k<k+\frac{\Delta}{2}} \tilde{\phi}(\vec{k},t)\tilde{\phi}(-\vec{k},t) ,
    \end{equation}
    %
    where \(k=|\vec{k}|\) denotes the modulus of \(\vec{k}\)
    and \(n_k\) is the number of lattice sites in a shell of radius \(k\)
    and thickness \(\Delta=2\pi/L_V\). The average domain size can then be
    defined in terms of the \(n\)-th moment of the structure factor
    \cite{laradji_molecular_1996}
    %
    \begin{equation}
        L_n(t) = 2\pi \left( \frac{\sum_k S(k,t)}{\sum_k k^n S(k,t)} \right)^{\frac{1}{n}}
        \label{eq:spherically_averaged_structure_factor}
    \end{equation}

@@

    %
    \begin{equation}
    L_\beta(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}k_\beta^2 S(\vec{k},t)}} .
    \end{equation}
    %
    This allows us to compute a domain size parallel
    (\(L_\parallel=L_z\)) and perpendicular (\(L_\perp=(L_x+L_y)/2\)) to the
    direction of \(\vec{B}\) separately. The average domain size \(L_d(t)\)
    is computed as \(L_d(t)=\sum_\beta L_\beta(t) / 3\).
    %
    \begin{equation}
    L_d(t)=2\pi\sqrt{\frac{\sum_{\vec{k}}S(\vec{k},t)}{\sum_{\vec{k}}\sum_\beta k_\beta^2 S(\vec{k},t)}} .
    \label{eq:directional_structure_factor}
    \end{equation}
    %

@@

% \section[Conclusions]{Conclusions\protect\footnote{This section is based on work published in the Royal Society of Chemistry's Soft Matter journal. It is reproduced from Ref.~\cite{karthikeyan_formation_2024} with permission from the Royal Society of Chemistry.}}
\section{Conclusions}
